---
title: "Spatial distribution of bovine tuberculosis in Minas Gerais under the PNCEBT, 2014 to 2018"
output:
  flexdashboard::flex_dashboard:
    theme: 
      version: 4
      bootswatch: minty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(flexdashboard)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```


```{r message=FALSE, warning=FALSE, include=FALSE}
## Leitura e preparção dos dados

require(readxl)
dados <- read_excel("DadosAnos.xlsx", na="NA")
tub_final <- read_excel("TubFinal.xlsx", na="NA")

require(sf)
Mg <- st_read("MG_Municipios_2019.shp")

SpMg <- merge(Mg, tub_final, by = "CD_MUN", all.x = TRUE)
```

2014-2018
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Cases of bTB - 2014-2018

```{r MapMinas}
require(ggplot2)
map <- ggplot(SpMg) + 
  geom_sf(aes(fill=CFTA),
          color = "black",
          linetype = 1,
          lwd = 0.1) +
  scale_fill_viridis_c(option = "B", direction=-1, alpha=1)
map + theme_bw()

```

Column {data-width=350}
-----------------------------------------------------------------------

### Total number of positive animals for bTB

```{r}
valueBox(sum(dados$Soma_f), color = "warning")
```

### Mean of cases of bTB per year

```{r}
require(dplyr)
require(plotly)
tub_g <- dados%>%
  group_by(Ano)%>%
  summarize(CFTA = round(mean(Soma_f),2))

g <- ggplot(tub_g, aes(x=Ano, y=CFTA)) +
  geom_point() + geom_line() + theme_bw()

ggplotly(g)
```

2014
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Cases of bTB in 2014

```{r Ano14}
## Dados por ano
tub14 <- filter(dados, Ano==2014)
colnames(tub14)[4] <- "tub_14"
colnames(tub14)[5] <- "sif_14"

tub15 <- filter(dados, Ano==2015)
colnames(tub15)[4] <- "tub_15"
colnames(tub15)[5] <- "sif_15"

tub16 <- filter(dados, Ano==2016)
colnames(tub16)[4] <- "tub_16"
colnames(tub16)[5] <- "sif_16"

tub17 <- filter(dados, Ano==2017)
colnames(tub17)[4] <- "tub_17"
colnames(tub17)[5] <- "sif_17"

tub18 <- filter(dados, Ano==2018)
colnames(tub18)[4] <- "tub_18"
colnames(tub18)[5] <- "sif_18"

## Concatenando as informações com o shape (por ano)

## 2014
SpMg14 <- merge(Mg, tub14, by = "CD_MUN", all.x = TRUE)

## 2015
SpMg15 <- merge(Mg, tub15, by = "CD_MUN", all.x = TRUE)

## 2016
SpMg16 <- merge(Mg, tub16, by = "CD_MUN", all.x = TRUE)

## 2017
SpMg17 <- merge(Mg, tub17, by = "CD_MUN", all.x = TRUE)

## 2018
SpMg18 <- merge(Mg, tub18, by = "CD_MUN", all.x = TRUE)

## Plot do mapa

map14 <- ggplot(SpMg14) + 
  geom_sf(aes(fill=tub_14),
          color = "black",
          linetype = 1,
          lwd = 0.1) +
  scale_fill_viridis_c(option = "B", direction=-1, alpha=1)
map14 + theme_bw()
```

Column {data-width=350}
-----------------------------------------------------------------------

### Total number of positive animals for bTB in 2014

```{r total}
valueBox(sum(tub14$tub_14), color = "warning")
```

### Difference between the number of animals with suggestive lesions and the number of positive animals - The 10 cities with the biggest discrepancy

```{r}
tub14 <- data.frame(tub14, "DIF"=tub14$tub_14-tub14$sif_14)

#Max_mun14 <- tub14[tub14$tub_14>31,c(3,4,5,11)]
Max_mun14 <- dplyr::arrange(tub14, DIF)[1:10,c(3,4,5,11)]
colnames(Max_mun14) <- c("Mun","CFTA","AIA","DIF")

DT::datatable(Max_mun14, rownames=F, options=list(paging=F))
```

2015
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Cases of bTB in 2015
```{r Ano2015}
map15 <- ggplot(SpMg15) + 
  geom_sf(aes(fill=tub_15),
          color = "black",
          linetype = 1,
          lwd = 0.1) +
  scale_fill_viridis_c(option = "B", direction=-1, alpha=1)
map15 + theme_bw()
```

Column {data-width=350}
-----------------------------------------------------------------------

### Total number of positive animals for bTB in 2015

```{r}
valueBox(sum(tub15$tub_15), color = "warning")
```

### Difference between the number of animals with suggestive lesions and the number of positive animals - The 10 cities with the biggest discrepancy

```{r}
tub15 <- data.frame(tub15, "DIF"=tub15$tub_15-tub15$sif_15)

#Max_mun15 <- tub15[tub15$tub_15>27,c(3,4,5)]
Max_mun15 <- dplyr::arrange(tub15, DIF)[1:10,c(3,4,5,11)]
colnames(Max_mun15) <- c("Mun","CFTA","AIA","DIF")

DT::datatable(Max_mun15, rownames=F, options=list(paging=F))
```

2016
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Cases of bTB in 2016
```{r}
map16 <- ggplot(SpMg16) + 
  geom_sf(aes(fill=tub_16),
          color = "black",
          linetype = 1,
          lwd = 0.1) +
  scale_fill_viridis_c(option = "B", direction=-1, alpha=1)
map16 + theme_bw()
```

Column {data-width=350}
-----------------------------------------------------------------------

### Total number of positive animals for bTB in 2016

```{r}
valueBox(sum(tub16$tub_16), color = "warning")
```

### Difference between the number of animals with suggestive lesions and the number of positive animals - The 10 cities with the biggest discrepancy

```{r}
tub16 <- data.frame(tub16, "DIF"=tub16$tub_16-tub16$sif_16)

Max_mun16 <- dplyr::arrange(tub16, DIF)[1:10,c(3,4,5,11)]
colnames(Max_mun16) <- c("Mun","CFTA","AIA","DIF")
#Max_mun16 <- tub16[tub16$tub_16>14,c(3,4,5)]

DT::datatable(Max_mun16, rownames=F, options=list(paging=F))
```

2017
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Cases of bTB in 2017
```{r}
map17 <- ggplot(SpMg17) + 
  geom_sf(aes(fill=tub_17),
          color = "black",
          linetype = 1,
          lwd = 0.1) +
  scale_fill_viridis_c(option = "B", direction=-1, alpha=1)
map17 + theme_bw()
```

Column {data-width=350}
-----------------------------------------------------------------------

### Total number of positive animals for bTB in 2017

```{r}
valueBox(sum(tub17$tub_17), color = "warning")
```

### Difference between the number of animals with suggestive lesions and the number of positive animals - The 10 cities with the biggest discrepancy

```{r}
tub17 <- data.frame(tub17, "DIF"=tub17$tub_17-tub17$sif_17)

Max_mun17 <- dplyr::arrange(tub17, DIF)[1:10,c(3,4,5,11)]
colnames(Max_mun17) <- c("Mun","CFTA","AIA","DIF")
#Max_mun17 <- tub17[tub17$tub_17>14,c(3,4,5)]

DT::datatable(Max_mun17, rownames=F, options=list(paging=F))
```

2018
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Cases of bTB in 2018
```{r}
map18 <- ggplot(SpMg18) + 
  geom_sf(aes(fill=tub_18),
          color = "black",
          linetype = 1,
          lwd = 0.1) +
  scale_fill_viridis_c(option = "B", direction=-1, alpha=1)
map18 + theme_bw()
```

Column {data-width=350}
-----------------------------------------------------------------------

### Total number of positive animals for bTB in 2018

```{r}
valueBox(sum(tub18$tub_18), color = "warning")
```

### Difference between the number of animals with suggestive lesions and the number of positive animals - The 10 cities with the biggest discrepancy

```{r}
tub18 <- data.frame(tub18, "DIF"=tub18$tub_18-tub18$sif_18)

Max_mun18 <- dplyr::arrange(tub18, DIF)[1:10,c(3,4,5,11)]
colnames(Max_mun18) <- c("Mun","CFTA","AIA","DIF")
#Max_mun18 <- tub18[tub18$tub_18>14,c(3,4,5)]

DT::datatable(Max_mun18, rownames=F, options=list(paging=F))
```

Descriptive measures
=======================================================================

Column
-----------------------------------------------------------------------

### Descriptive statistics (2008-2018)
```{r EstDesc}
desc1 <- fBasics::basicStats(tub_final$CFTA)
desc2 <- fBasics::basicStats(tub_final$AIA)[[1]]
desc3 <- fBasics::basicStats(tub_final$PPB)[[1]]
desc4 <- fBasics::basicStats(tub_final$PER)[[1]]

EstDesc <- data.frame("CFTA"=desc1,"AL"=desc2,"PPB"=desc3,"PER"=desc4)
colnames(EstDesc)[1] <- "CFTA"

DT::datatable(EstDesc, filter = "none", rownames=T, options=list(paging=F))
```

Column
-----------------------------------------------------------------------

### Difference between the number of animals with suggestive lesions and the number of positive animals - The 10 cities with the biggest discrepancy (2014 to 2018)
```{r}
## Diferença negativa (10 mais diferentes negativamente)
dados2 <- data.frame(dados, "Dif"=dados$Soma_f-dados$SIF)
mun1 <- filter(dados2, Dif==-107) #Uberaba
mun2 <- filter(dados2, Dif==-85) #Unaí
mun3 <- filter(dados2, Dif==-77) #Araxá
mun4 <- filter(dados2, Dif==-51) #Patrocínio 2014
mun5 <- filter(dados2, Dif==-48) #Patrocínio 2017
mun6 <- filter(dados2, Dif==-44) #Patrocínio 2016 e Serra do Salitre 2015
mun7 <- filter(dados2, Dif==-43) #Campo Florido 2015 Perdizes 2014
mun8 <- filter(dados2, Dif==-40) #Patos de Minas

Mun_neg <- data.frame(rbind(mun1,mun2,mun3,mun4,mun5,mun6,mun7,mun8))
Mun_neg <- Mun_neg[,c(1:8,11)]
colnames(Mun_neg) <- c("CD_MUN","ANO","MUN","CFTA","AL","FC","CH","MC","DIF")
DT::datatable(Mun_neg, rownames=F, options=list(paging=F))
```

<!--### Municípios com maior número de bovinos com bTB segundo ano de notificação.
```{r}
A14 <- filter(tub14, tub_14==max(tub14$tub_14, na.rm = T))[,2:5]
colnames(A14)[3:4] <- c("CFTA","AIA")

A15 <- filter(tub15, tub_15==max(tub15$tub_15, na.rm = T))[,2:5]
colnames(A15)[3:4] <- c("CFTA","AIA")

A16 <- filter(tub16, tub_16==max(tub16$tub_16, na.rm = T))[,2:5]
colnames(A16)[3:4] <- c("CFTA","AIA")

A17 <- filter(tub17, tub_17==max(tub17$tub_17, na.rm = T))[,2:5]
colnames(A17)[3:4] <- c("CFTA","AIA")

A18 <- filter(tub18, tub_18==max(tub18$tub_18, na.rm = T))[,2:5]
colnames(A18)[3:4] <- c("CFTA","AIA")

Max_mun <- data.frame(rbind(A14,A15,A16,A17,A18))

DT::datatable(Max_mun, rownames=F, options=list(paging=F))
```

-->